---
title: "The Usual Suspects"
author: "Jack Carter"  
date: "8/15/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(stringr)
library(ggthemes)
library(stats)
library(factoextra)

# sets the seed for random number generation. 
set.seed(1995)

# sets max overlaps in GGplot.  
options(ggrepel.max.overlaps = 15)

# my personal plot theme for data visualizations. 
my_theme <- theme_economist_white(gray_bg = FALSE) +
  theme(plot.title = element_text(hjust = 0.5,
                                  vjust = 10,
                                  size = 10,
                                  color = "#474747"),
        plot.margin = unit(c(1, 1, 1, 1), "cm"),
        axis.text = element_text(size = 9,
                                 color = "gray30"),
        axis.text.x=element_text(vjust = -2.5),
        axis.title.x = element_text(size = 9,
                                    color = "gray30",
                                    vjust = -10),
        axis.title.y = element_text(size = 9,
                                    color = "gray30",
                                    vjust = 10),
        legend.direction = "vertical", 
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size = 11,
                                   color = "gray20"),
        legend.margin=margin(1, -15, 1, 0),
        legend.spacing.x = unit(0.25, "cm"),
        legend.key.size = unit(1, "cm"), 
        legend.key.height = unit(0.75, "cm"),
        strip.text = element_text(hjust = 0.5,
                                  vjust = 1,
                                  size = 10,
                                  color = "#474747"),
        panel.spacing = unit(2, "lines"))

# sets the working directory. 
setwd("C:/Users/HUAWEI/Desktop/Exploratory data/The Magnificent Seven")

# reads in the data. 
data <- read_csv("Data/combined data/clean_data.csv")

# sets a countries variable. 
countries <- tibble(countries = data$country)

# creates a data frame of country name reassignments. 
reassignments <- tibble(original = c("Bosnia and Herzegovina",
                                     "Congo, Dem. Rep.",
                                     "Congo, Rep.",
                                     "Egypt, Arab Rep.",
                                     "Gambia, The",
                                     "Korea, Rep.",
                                     "Kyrgyz Republic",
                                     "Lao PDR",
                                     "North Macedonia",
                                     "Russian Federation",
                                     "Sao Tome and Principe",
                                     "Slovak Republic",
                                     "South Sudan",
                                     "Syrian Arab Republic",
                                     "United Arab Emirates",
                                     "United Kingdom",
                                     "United States",
                                     "Venezuela, RB",
                                     "West Bank and Gaza",
                                     "Yemen, Rep.",
                                     "Hong Kong SAR, China",
                                     "Iran, Islamic Rep."),
                        new = c("Bosnia",
                                "DRC",
                                "Congo",
                                "Egypt",
                                "Gambia",
                                "S. Korea",
                                "Kyrgyzstan",
                                "Laos",
                                "N. Macedonia",
                                "Russia",
                                "Sao Tome",
                                "Slovakia",
                                "S. Sudan",
                                "Syria",
                                "UAE",
                                "UK",
                                "US",
                                "Venezuela",
                                "Palestine",
                                "Yemen",
                                "Hong Kong",
                                "Iran"))

# reassigns the country names for all data frames. 
reassign_countries <- function(data) {
  countries <- data
  for(i in 1:length(countries)) {
    countries[i] <- if(countries[i] %in% reassignments$original) {
      reassignments$new[which(countries[i]==reassignments$original)]
    }
    else {
      countries[i]
    }
  }
  return(countries)
}

# reassigns all countries in all data frames. 
for(i in 1:length(countries)) {
  countries$countries <- reassign_countries(countries$countries)
}

# takes the countries column away from the data frame. 
data <- data[-1]

# sets the row names to the country names. 
rownames(data) <- countries$countries

# creates a k-means object. 
kmeans.res <- kmeans(data, centers=3, iter.max = 15, nstart = 1)

# creates a list of colors. 
green <- which(table(kmeans.res$cluster)==42)
orange <- which(table(kmeans.res$cluster)==75)
red <- which(table(kmeans.res$cluster)==48)

col_values <- c("#009900", "#FF8000", "#CC0000")
cols <- c(col_values[green], col_values[orange], col_values[red])

```

## **Summary**

Based on eight different development indicators, the world's countries can be grouped into three categories. While some previously underdeveloped countries such as Singapore have shot into the so-called "first world", many of the group's occupants remain the usual suspects. 

&nbsp;

```{r, echo = FALSE, message = FALSE, warning = FALSE, dpi=600}

# visualizes the data with the factoextra package. 
fviz_cluster(object=kmeans.res,
             data=data,
             stand=FALSE,
             geom = "text",
             repel=TRUE,
             ellipse.type = "t",
             ellipse.level=0.95,
             ellipse.alpha = 0.1,
             labelsize = 5) +
  scale_colour_manual(values = cols) +
  scale_fill_manual(values = cols) + 
  ggtitle("International Development") +
  my_theme +
  theme(legend.position="none")

```

&nbsp;

## **Disclaimer**

Thinking of countries only in terms of these three groups obscures the fact that development is actually a continuum. Also, while many African countries may remain in the so-called "third world", significant progress has been achieved here in recent decades. 

&nbsp;

## **Method**

### **1) Reconcile Data Sets:**

Data from the World Bank, Transparency International, Worldwide Governance Indicators and the UN are reconciled by identifying, changing and dropping relevant country names.

---EXAMPLE CODE SNIPET---

```{r, echo = TRUE, message = FALSE, warning = FALSE, dpi=600}

# creates a list of non-identical country names. 
get_non_identical <- function(list_1, list_2) {
  data <- list_2[which(list_2 %in% list_1 == FALSE)]
  return(data)
}

```

&nbsp;

### **2) Create Padded NA values:**

Padded NA values are created for each country which is in the World Bank data frame but not in other data frames. 

---EXAMPLE CODE SNIPET---

```{r, echo = TRUE, message = FALSE, warning = FALSE, dpi=600}

# creates a data frame of zeros for missing countries. 
create_padded_values <- function(data_frame) {
  # sets the column names with a value of NA. 
  cols <- data_frame[1,-1]
  for(i in 1:ncol(cols)) {
    cols[1,i] <- NA
  }
  # creates a list of missing countries with col values of NA. 
  added_countries <- list()
  for(i in 1:length(missing)) {
    added_countries[[i]] <- if(!missing[i] %in% data_frame$country) {
      tibble(country = missing[i]) %>%
        cbind(cols) %>%
        tibble()
    }
  }
  # combines all countries into 1 data frame if there are any missing. 
  if(length(added_countries)>0) {
    df <- added_countries %>% 
      reduce(rbind)
  } else {
    return(tibble(country = "No missing countries here!"))
  }
  return(df)
}

```

&nbsp;

### **3) Remove High NA Entries:**

Countries with a high number of NAs across multiple measures are identified and removed. 

---EXAMPLE CODE SNIPET---

```{r, echo = TRUE, message = FALSE, warning = FALSE, dpi=600}

# gets how many NAs a certain row has. 
get_nas <- function(row) {
  nas <- list()
  for(i in 1:ncol(row)) {
    nas[i] <- if(is.na(row[,i])) 1 else 0
  }
  return(sum(unlist(nas)))
}

```

&nbsp;

### **4) Check Outliers**

Outliers are checked for and capped by determining whether any of the values lie further than the interquartile range (IQR) * 1.5 away from the mean. 

```{r, echo = TRUE, message = FALSE, warning = FALSE, dpi=600}

# checks each column for outliers and caps them at IQR * 1.5. 
cap_col_outliers <- function(column) {
  iqr <- IQR(data[,2][!is.na(data[,2])])
  upper_quantile <- quantile(data[,2][!is.na(data[,2])], 3/4, names=FALSE)
  lower_quantile <- quantile(data[,2][!is.na(data[,2])], 1/4, names=FALSE)
  
  high <- upper_quantile + (iqr*1.5)
  low <- lower_quantile - (iqr*1.5)
  
  outliers <- list()
  for(i in 1:nrow(column)) {
    value <- pull(column[i,])
    outliers[[i]] <- if(!is.na(value) & value > low & value < high) {
      value
    } else {
      if(!is.na(value) & value < low) {
        low
      } else {
        if(!is.na(value) & value > high) {
          high
        } else {
          NA
        }
      }
    }
  }
  return(unlist(outliers))
}

```

&nbsp;

### **5) Get Z-scores**

The data is standardized by calculating the z-scores as the data point minus the mean, divided by the standard deviation. 

```{r, echo = TRUE, message = FALSE, warning = FALSE, dpi=600}

# gets the z scores for individual columns. 
get_col_z_scores <- function(column) {
  avg <- mean(pull(column[!is.na(column),]))
  std <- sd(pull(column[!is.na(column),]))
  
  z_scores <- list()
  for(i in 1:nrow(column)) {
    z_scores[[i]] <- (pull(column[i,]) - avg) / std
  }
  return(unlist(z_scores))
}

```

&nbsp;

### **6) Create Indexes**

The individual data columns are then combined into 8 indexes, including technology, politics, education, health, happiness, economic, equality, poverty. 

```{r, echo = TRUE, message = FALSE, warning = FALSE, dpi=600}

# gets a new index measure by averaging column scores. 
create_index <- function(data, col_names) {
  measures <- data[,col_names]
  cols <- length(col_names)
  index <- list() 
  for(i in 1:nrow(measures)) {
    num <- sum(!is.na(measures[i,c(1:cols)]))
    index[[i]] <- if(sum(is.na(measures[i,c(1:cols)])) >= cols) {
      NA
    } else {
      sum(measures[i,which(!is.na(measures[i,c(1:cols)]))]) / num
    }
  }
  unlist(index)
}

```

&nbsp;

### **7) Impute NAs**

The remaining NA values are then imputed by taking a grouped average based on a complete column with the most variability.  

```{r, echo = TRUE, message = FALSE, warning = FALSE, dpi=600}

# imputes missing values by taking a grouped average based on a complete 
# column we have full data for with the most variability.  
fill_NAs <- function(data, measure) {
  impute_var <- get_best_impute_var(data)
  
  # gets the grouped averages for missing values. 
  imputed_values <- data %>%
    mutate(impute_var = round(eval(parse(text=impute_var)))) %>%
    group_by(impute_var) %>%
    filter(!is.na(eval(parse(text=measure)))) %>%
    summarise(avg = mean(eval(parse(text=measure))))
  
  # imputes the missing values for the target column. 
  NAs <- data[is.na(data[,measure]),]
  imputations <- list()
  for(i in 1:nrow(NAs)) {
    imputations[[i]] <- imputed_values$avg[which(pull(imputed_values["impute_var"]) == pull(round(NAs[i,impute_var])))]
  }
  data[is.na(data[,measure]),measure] <- unlist(imputations)
  return(data[,measure])
}

```

&nbsp;

### **8) Visualize Data**

The data is visualized with the use of a K-means clustering algorithm and the factoextra visualization package

```{r, echo = TRUE, message = FALSE, warning = FALSE, dpi=600}

# visualizes the data with the factoextra package. 
plot <- fviz_cluster(object=kmeans.res,
             data=data,
             stand=FALSE,
             geom = "text",
             repel=TRUE,
             ellipse.type = "t",
             ellipse.level=0.95,
             ellipse.alpha = 0.1,
             labelsize = 8) +
  scale_colour_manual(values = cols) +
  scale_fill_manual(values = cols) + 
  ggtitle("International Development") +
  my_theme +
  theme(legend.position="none")

```

&nbsp;

## **Sources**

- Transaprency International (2022) https://www.transparency.org/en/

- WGI (2022) http://info.worldbank.org/governance/wgi/

- World Bank (2022) https://data.worldbank.org/

- World Happiness Report (2022) https://worldhappiness.report/ed/2022/ 

&nbsp;
&nbsp;
&nbsp;
